---
title: "COVID-19 Exploration"
author: "Dr. Clifton Franklund"
date: "4/18/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

## The data used for this document

| Variable   | Description                                      | Source |
|:----------:|:------------------------------------------------:|:------:|
| State      | State names, alphabetical                        | By hand |
| Governor   | Political party affiliation of current governor  | [Wikipedia](https://en.wikipedia.org/wiki/List_of_United_States_governors) |
| Tests      | Estimated number of COVID-19 tests per state     | [Politico.com](https://www.politico.com/interactives/2020/coronavirus-testing-by-state-chart-of-new-cases/) |
| Cases      | Estimated number of COVID-19 cases per state     | [Wikipedia](https://en.wikipedia.org/wiki/2020_coronavirus_pandemic_in_the_United_States) |
| Deaths     | Estimated number of COVID-19 deaths per state    | [Wikipedia](https://en.wikipedia.org/wiki/2020_coronavirus_pandemic_in_the_United_States) |
| Population | Estimated 2019 census value for state population | [Wikipedia](https://en.wikipedia.org/wiki/List_of_states_and_territories_of_the_United_States_by_population) |
| Area       | State area expressed in square miles             | [Statesymbolusa.org](https://statesymbolsusa.org/symbol-official-item/national-us/uncategorized/states-size) |



```{r data}
# Loading data from Excel spreadsheet with web-derived values
myData <- read_xlsx("COVID19_data.xlsx")

# Calculating average population density per state (people per square mile)
myData$Population.Density <- myData$Population/myData$Area

# Calculating normalized testing value (tests per million people)
myData$Tests.Norm <- myData$Tests/myData$Population*1000000

# Calculating normalized cases (cases per million people)
myData$Cases.Norm <- myData$Cases/myData$Population*1000000

# Calculating normalized deaths (deaths per million people)
myData$Death.Norm <- myData$Deaths/myData$Population*1000000

```

## Including Plots
